---
title: "AUC_clinChem"
author: "Kristen James"
date: "6/1/2022"
output: html_document
---

6/1/22

Prep the BD1-4 clinical chemistry data. Calculate AUC for each blood marker. 

Trapezoid rule:
A = ((a + b)/2)h

a = TMAO at time X1
b = TMAT at time X2
h = time X2 - time X1

9/12/22

Update - add HDL, LDL, total cholesterol, TNF-a, IL6, CRP to analysis


```{r libraries, results='hide', message=FALSE}
library(dplyr)
```

# Load 

```{r load biocrates data, results='hide'}
# Load data from REDCap
getwd()
clin <- read.csv("../../Data/Raw/REDCap/CTSC24532USDAWHNRCNu-GlucoseInsulinTrigly_DATA_2022-06-01_1224.csv", na.strings=c(""," ","NA"), row.names = 1)

head(clin)
```

# Glucose

```{r AUC - Glucose}
# Specify correct blood marker
phen_vars <- c("glc_bd1", "glc_bd2", "glc_bd3", "glc_bd4")
pheno <- "glucose"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- clin[,colnames(clin) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr)
shapiro.test(log(phen_df$AUC_pheno_hr))
shapiro.test((phen_df$AUC_pheno_hr)^.5)
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # best transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$glucose_AUC_pheno_hr)

# Save
glucose_df_noNA <- phen_df_noNA
```

# Insulin

```{r AUC - Insulin}
# Specify correct blood marker
phen_vars <- c("insulin_bd1", "insulin_bd2", "insulin_bd3", "insulin_bd4")
pheno <- "insulin"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- clin[,colnames(clin) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr)
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$insulin_AUC_pheno_hr)

# Save
insulin_df_noNA <- phen_df_noNA
```

# TG

```{r AUC - TG}
# Specify correct blood marker
phen_vars <- c("tg_bd1", "tg_bd2", "tg_bd3", "tg_bd4")
pheno <- "triglycerides"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- clin[,colnames(clin) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr)
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$triglycerides_AUC_pheno_hr)

# Save
triglycerides_df_noNA <- phen_df_noNA
```

9/12/22 Add other phenotypes

Load other data

```{r Load lipid data}
lipid <- read.csv("../../Data/Raw/REDCap/CTSC24532USDAWHNRCNu-LipidData_DATA_2022-06-01_1213.csv", row.names = 1)
colnames(lipid)
```


# HDL

```{r AUC - HDL}
# Specify correct blood marker
phen_vars <- c("hdl_bd1", "hdl_bd2", "hdl_bd3", "hdl_bd4")
pheno <- "HDL"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- lipid[,colnames(lipid) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr)
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$HDL_AUC_pheno_hr)

# Save
HDL_df_noNA <- phen_df_noNA
```

# LDL

```{r AUC - LDL}
# Specify correct blood marker
phen_vars <- c("ldl_bd1", "ldl_bd2", "ldl_bd3", "ldl_bd4")
pheno <- "LDL"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- lipid[,colnames(lipid) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr) # fine
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$LDL_AUC_pheno_hr)

# Save
LDL_df_noNA <- phen_df_noNA
```

# Total Cholesterol

```{r AUC - Total Cholesterol}
# Specify correct blood marker
phen_vars <- c("chol_bd1", "chol_bd2", "chol_bd3", "chol_bd4")
pheno <- "chol"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- lipid[,colnames(lipid) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd2_30m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd2_30m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-30 min
phen_df$area_0_30 <- (((phen_df$bd1_0m + phen_df$bd2_30m)/2) * 30)

# BD 30-180 min (150 min)
phen_df$area_30_180 <- (((phen_df$bd2_30m + phen_df$bd3_180m)/2) * 150)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_30 + phen_df$area_30_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr) # fine
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$chol_AUC_pheno_hr)

# Save
chol_df_noNA <- phen_df_noNA
```

# Load inflammatory data

No BD2 for these variables.

```{r Load inflammatory data}
infl <- read.csv("../../Data/Raw/REDCap/CTSC24532USDAWHNRCNu-CytokinesChemokines_DATA_2022-09-12_1028.csv", row.names = 1)
colnames(infl)
```

# TNFa 

No BD2 for these variables.

```{r AUC - tnfa}
# Specify correct blood marker
phen_vars <- c("tnfa_bd1", "tnfa_bd3", "tnfa_bd4")
pheno <- "tnfa"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- infl[,colnames(infl) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-180 min
phen_df$area_0_180 <- (((phen_df$bd1_0m + phen_df$bd3_180m)/2) * 180)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr) 
shapiro.test(log(phen_df$AUC_pheno_hr)) # best transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # works transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$tnfa_AUC_pheno_hr)

# Save
tnfa_df_noNA <- phen_df_noNA
```

# IL6

No BD2 for these variables.

```{r AUC - IL6}
# Specify correct blood marker
phen_vars <- c("il6_bd1", "il6_bd3", "il6_bd4")
pheno <- "il6"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- infl[,colnames(infl) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-180 min
phen_df$area_0_180 <- (((phen_df$bd1_0m + phen_df$bd3_180m)/2) * 180)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr) 
shapiro.test(log(phen_df$AUC_pheno_hr)) # works transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # best transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$il6_AUC_pheno_hr)

# Save
il6_df_noNA <- phen_df_noNA
```

# CRP

No BD2 for these variables.

```{r AUC - CRP}
# Specify correct blood marker
phen_vars <- c("crp_bd1", "crp_bd3", "crp_bd4")
pheno <- "crp"

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Run Chunk
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Can chunk this and run
# Subset
phen_df <- infl[,colnames(infl) %in% phen_vars]
# Rename columns to make code recyclable 
colnames(phen_df) <- c("bd1_0m", "bd3_180m", "bd4_360m")

# Make sure there are no values below 0. Not only would that not make physiological sense but we'd need to subtract the area. 
summary(phen_df$bd1_0m)
summary(phen_df$bd3_180m)
summary(phen_df$bd4_360m)

# BD 0-180 min
phen_df$area_0_180 <- (((phen_df$bd1_0m + phen_df$bd3_180m)/2) * 180)

# BD 180-360 min (180 min)
phen_df$area_180_360 <- (((phen_df$bd3_180m + phen_df$bd4_360m)/2) * 180)

# Total area
phen_df$AUC_pheno_min <- phen_df$area_0_180 + phen_df$area_180_360

# Check it out. The unit is uM-min.
summary(phen_df$AUC_pheno_min)

# Make a uM-hour variable
phen_df$AUC_pheno_hr <- (phen_df$AUC_pheno_min/60) 

# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Manually
# ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

# Manually,
# Check out the distribution
shapiro.test(phen_df$AUC_pheno_hr) 
shapiro.test(log(phen_df$AUC_pheno_hr)) # works transformation
shapiro.test((phen_df$AUC_pheno_hr)^-.5) # best transformation

# Transform
phen_df$AUC_pheno_hr_NegSqrRt <- phen_df$AUC_pheno_hr ^ -.5
phen_df$AUC_pheno_min_NegSqrRt <- phen_df$AUC_pheno_min ^ -.5
phen_df$AUC_pheno_hr_ln <- log(phen_df$AUC_pheno_hr)
phen_df$AUC_pheno_min_ln <- log(phen_df$AUC_pheno_min)

# Rename columns to fit variable
colnames(phen_df)
colnames(phen_df) <- paste0(pheno, "_", colnames(phen_df))
colnames(phen_df)

# Complete cases
phen_df_noNA <- na.omit(phen_df)

# Plot
plot(phen_df_noNA$crp_AUC_pheno_hr)

# Save
crp_df_noNA <- phen_df_noNA
```

# Merge data frames

```{r}
df <- merge(triglycerides_df_noNA, glucose_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, insulin_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, HDL_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, LDL_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, chol_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, tnfa_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, il6_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]

df <- merge(df, crp_df_noNA, by = 0)
rownames(df) <- df[,1]
df <- df[,-1]
```

Save

```{r Save}
# 6/11/2022
#write.csv(df, "../../Data/Processed/Clinical_AreaUnderCurve.csv")
#saveRDS(df, "../../Data/Processed/Clinical_AreaUnderCurve.rds")

# 9/12/2022
write.csv(df, "../../Data/Processed/Clinical_Inflammation_AreaUnderCurve.csv")
saveRDS(df, "../../Data/Processed/Clinical_Inflammation_AreaUnderCurve.rds")
```

Combine with TMAO data for quick correlations/t-tests

```{r}
tmao <- readRDS("../../Data/Processed/Biocrates_AreaUnderCurve_meta.rds")
rownames(tmao) <- tmao[,1]
tmao <- tmao[,-1]

dftmao <- merge(df, tmao, by=0, all.x=F, all.y = T)
rownames(dftmao) <- dftmao[,1]
dftmao <- dftmao[,-1]
```

Stats

```{r}
# AUC Totals
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ insulin_AUC_pheno_hr_NegSqrRt + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ triglycerides_AUC_pheno_hr_NegSqrRt + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ glucose_AUC_pheno_hr + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ HDL_AUC_pheno_hr_NegSqrRt + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ LDL_AUC_pheno_hr + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ chol_AUC_pheno_hr + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ tnfa_AUC_pheno_hr_ln + sex*age, dftmao)) # yes
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ il6_AUC_pheno_hr_NegSqrRt + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ crp_AUC_pheno_hr_NegSqrRt + sex*age, dftmao)) 

# 0
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ insulin_bd1_0m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ triglycerides_bd1_0m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ glucose_bd1_0m + sex*age, dftmao))

# 30
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ insulin_bd2_30m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ triglycerides_bd2_30m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ glucose_bd2_30m + sex*age, dftmao))

# 180
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ insulin_bd3_180m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ triglycerides_bd3_180m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ glucose_bd3_180m + sex*age, dftmao))

# 360
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ insulin_bd4_360m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ triglycerides_bd4_360m + sex*age, dftmao))
summary(lm(AUC_TMAO_uM_hour_NegSqrRt ~ glucose_bd4_360m + sex*age, dftmao))


#library(Hmisc)
#rcorr(x, type="pearson")
```

Maybe something related to TG

