---
title: "Preprocess Biocrates Data"
author: "Kristen James"
date: "2/10/2022"
html_document:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    keep_md: true
---

# Overview 

Project: FL100

Date: 2/10/2022

Purpose: Preliminary metabolomics data generated using Biocrates plates was shared by the Newman lab. The biocrates data file needs to be manipulated/cleaned so that it is ready to be analyzed in R.

Note, these are the metabolites quantified via biocrates, https://biocrates.com/wp-content/uploads/2021/01/biocrates-Quant500-list-of-metabolites-v4-2021.pdf 

```{r libraries, results='hide', message=FALSE}
library(dplyr)
```

# Load 

```{r load biocrates data, results='hide'}
# Load data from Jon Newman on 8/5/21
bio <- read.csv("../../Data/Raw/from_JohnNewman/Prelim_Biocretes_dataset_v3_corrected_PPTime.csv", na.strings=c(""," ","NA"))

head(bio)
```

# Clean 

This file is ugly. It needs to be tailored before we can confidently maneuver in it. 

```{r Beautify biocrates file, results='hide'}
# Set column names
colnames(bio) <- bio[3,]

# Note, can't set row names to subject ID yet bc this data-set has multiple blood draws/person so IDs aren't unique. 
# Will subset to BD1 in our first pass. See chunk below.

# Trim junk rows on top and NA rows on bottom
dim(bio) #683 591
bio <- bio[4:677,] #674 591

# Use gsub to get rid of spaces in column names - YUCK!
colnames(bio)[1:35]
colnames(bio) <- gsub(" ", "_", colnames(bio))

# Get rid of blank columns
emptycols <- sapply(bio, function (k) all(is.na(k)))
table(emptycols) # no empty cols now
bio2 <- bio[!emptycols]
dim(bio2) #674 591

# Look at structure
str(bio2) # Imported as chr. Not what we want. Check that out.
colnames(bio2)[1:35]
# First, subset to the columns we want to turn into numeric values
bio3 <- bio2[,c(9:length(bio2))]
# Then, lapply the combined character->numeric function to turn them into numbers
bio4 <- lapply(bio3, function(x) as.numeric(as.character(x)))
str(bio4)[1:35]
# Finally, turn your list into a data frame using cbind
bio5 <- do.call(cbind.data.frame, bio4)

# Combine with columns 1-8
bio6 <- cbind(bio2[,1:8], bio5)

# Last touches
bio6$Blood_Draw <- as.factor(bio6$Blood_Draw)
bio6$PlateNumber <- as.factor(bio6$PlateNumber)
bio6$Subject_ID <- as.factor(bio6$Subject_ID)
bio6$pp_Time_min <- as.factor(bio6$pp_Time_min)
bio6$import_order <- as.numeric(bio6$import_order)
```

Save long data
```{r Save long biocrates data}
write.csv(bio6, "../../Data/Processed/Biocrates_long.csv")
saveRDS(bio6, "../../Data/Processed/Biocrates_long.rds")
```

# Combine with Metadata 

Combine with metadata such as sex, age, BMI, etc. 

```{r load TMAO, results='hide'}
# load TMAO
tmao <- read.csv("../../Data/Raw/TMAO/CTSC24532USDAWHNRCNu-TMAOKJ012220_DATA_2020-07-15_0915_NAedited.csv")

# Check it out and simplify
colnames(tmao)
tmao <- subset(tmao, select=c(subject_id, age, sex, bmi_final, bin_number, endo_ln_rhi, endo_al, cystatinc_bd1, tmao, betaine, choline, carnitine, phosphocholine, creatinine, tnfa_bd1))

# Make logTMAO column
tmao$lnTMAO_WangMethod <- log(tmao$tmao)

# For merging, check str of subject id
str(tmao$subject_id)
tmao$subject_id <- as.factor(tmao$subject_id)

```

Left add
```{r left_join biocrates and meta data}
bio6_tmao <- left_join(bio6, tmao, by = c("Subject_ID"="subject_id"))
```

Save long data with meta
```{r Save long biocrates data}
write.csv(bio6_tmao, "../../Data/Processed/Biocrates_long_meta.csv")
saveRDS(bio6_tmao, "../../Data/Processed/Biocrates_long_meta.rds")
```

# Subset by Blood Draw

Subset to each blood draw samples and set row names.
```{r Subset by blood draw, results='hide'}
# Subset to BD
table(bio6_tmao$pp_Time_min)
str(bio6_tmao$pp_Time_min)

# Make dfs by BD
bio_bd1 <- bio6_tmao[bio6_tmao$pp_Time_min == 0,]
bio_bd2 <- bio6_tmao[bio6_tmao$pp_Time_min == 30,]
bio_bd3 <- bio6_tmao[bio6_tmao$pp_Time_min == 180,]
bio_bd4 <- bio6_tmao[bio6_tmao$pp_Time_min == 360,]

# Set row names
# Now set row names to subject ID
rownames(bio_bd1) <- bio_bd1[,"Subject_ID"]
rownames(bio_bd2) <- bio_bd2[,"Subject_ID"]
rownames(bio_bd3) <- bio_bd3[,"Subject_ID"]
rownames(bio_bd4) <- bio_bd4[,"Subject_ID"]
```

Save time course data for future use
```{r Save Biocrates data}
# BD1
write.csv(bio_bd1, "../../Data/Processed/Biocrates_BD1_meta.csv")
saveRDS(bio_bd1, "../../Data/Processed/Biocrates_BD1_meta.rds")

# BD2
write.csv(bio_bd2, "../../Data/Processed/Biocrates_BD2_meta.csv")
saveRDS(bio_bd2, "../../Data/Processed/Biocrates_BD2_meta.rds")

# BD3
write.csv(bio_bd3, "../../Data/Processed/Biocrates_BD3_meta.csv")
saveRDS(bio_bd3, "../../Data/Processed/Biocrates_BD3_meta.rds")

# BD4
write.csv(bio_bd4, "../../Data/Processed/Biocrates_BD4_meta.csv")
saveRDS(bio_bd4, "../../Data/Processed/Biocrates_BD4_meta.rds")

```

