---
title: "Biocrates_Microbiome"
author: "Kristen James"
date: "6/9/2022"
output: html_document
---

Does peak-TMAO group relate to the gut microbiome?

To answer this, load the phyloseq (PS) object and alpha-stat data used in FL100_TMAO. Modify the PS object to include biocrates data of interest. See if alpha and beta diversity differ by peak TMAO group. Get peak TMAO group info from TMAO_Meal_Response.Rmd file. 

```{r set up}
library(phyloseq)
# Clear environment to start fresh (if necessary)
rm(list = ls())

# PSO
load(file = "../../Data/from_FL100_TMAO/PSOtmao1G_211118.RData") #PSOtmao1G
#PSO <- PSOtmao1G # shorter name

# Get sample data which includes alpha stats
df <- as(sample_data(PSOtmao1G), "data.frame")
colnames(df)

# Get biocrates data and peak group
#phen <- readRDS("../../Data/Processed/dfwide_metavariables_n97.rds")
phen <- read.csv("../../Data/Processed/dfwide_metavariables_n97_inflam_eth.csv", row.names = 1)
colnames(phen)
# Try not to duplicate columns 
removeCol <- c("age", "sex", "bmi_final", "age_cat", "bmi_cat", "bmi_final","tmao", "choline", "betaine", "carnitine", "phosphocholine", "creatinine", "endo_ln_rhi")
phen <- phen[, !(colnames(phen) %in% removeCol)] # subject-ID is not needed


# Merge
df2 <- merge(phen, df, by = 0)
row.names(df2) <- df2[,"Row.names"]
df2 <- subset(df2, select =-c(Row.names))

colnames(df2)
```

```{r alpha diversity}
#~~~~~~~~~~~~~~~~
# Compute
#~~~~~~~~~~~~~~~~
# Check distribution
shapiro.test(df2$shannon) # not normal, proceed with caution
shapiro.test(log(df2$shannon))
shapiro.test((df2$shannon)^.5)

shapiro.test(df2$faith_pd) # normal
shapiro.test(df2$observed_otus) # normal
shapiro.test(log(df2$pielou_e))
shapiro.test((df2$pielou_e)^.5) # not normal

# Run linear regression looking at potential covariates and the different blood draws
summary(lm(tmao_log ~ shannon + sex*age + cystatinc_bd1 + bmi_final.x, df2))
summary(lm(log(TMAO.0) ~ shannon + sex*age + cystatinc_bd1 + bmi_final.x, df2))
summary(lm(log(TMAO.30) ~ shannon + sex*age + cystatinc_bd1 + bmi_final.x, df2))
summary(lm(log(TMAO.180) ~ shannon + sex*age + cystatinc_bd1 + bmi_final.x, df2))
summary(lm(log(TMAO.360) ~ shannon + sex*age + cystatinc_bd1 + bmi_final.x, df2))
# For manuscript
summary(lm(log(TMAO.0) ~ shannon + sex*age, df2))

# Run linear regression, bare bones
summary(lm(shannon ~ peakTMAO, df2))
summary(lm(faith_pd ~ peakTMAO, df2))
summary(lm(observed_otus ~ peakTMAO, df2))
summary(lm(pielou_e ~ peakTMAO, df2))

# Run linear regression, no evidence to include BMI
summary(lm(log(AUC_TMAO_uM_hour) ~ shannon + sex*age + cystatinc_bd1, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ faith_pd + sex*age + cystatinc_bd1, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ observed_otus + sex*age + cystatinc_bd1, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ pielou_e + sex*age + cystatinc_bd1, df2))

# Cystatin C is not related to AUC-TMAO. Remove it from the model.
# Report these stats
summary(lm(log(AUC_TMAO_uM_hour) ~ shannon + sex*age, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ faith_pd + sex*age, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ observed_otus + sex*age, df2))
summary(lm(log(AUC_TMAO_uM_hour) ~ pielou_e + sex*age, df2))

plot(df2$AUC_TMAO_uM_hour, df2$shannon)

# Check the residuals 
resid <- lm(log(AUC_TMAO_uM_hour) ~ shannon + sex*age, df2)
plot(resid$residuals) # they look random
summary(resid)
```
So alpha diversity is related to fasting values but not AUC-values. 

Scatter plot of alpha diversity measures by AUC

```{r shannon plot }
summary(lm(AUC_TMAO_uM_min_NegSqrRt ~ shannon + sex*age + cystatinc_bd1, df2))

x_variable <- "Shannon Statistic"
y_variable <- "Ln of Area Under the Curve TMAO"
#mdn_cut <- log(median(df2$AUC_TMAO_uM_min_NegSqrRt))
# Plot
shannon_plot <- ggplot(df2, aes(y=log(AUC_TMAO_uM_hour), x=shannon)) +
  geom_point(aes(shape=bmi_cat, color = age_cat)) +
  #geom_vline(xintercept = mdn_cut, linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, fullrange = F) +
  theme_bw() + 
  ylab(y_variable) +
  xlab(x_variable) +
  scale_shape_discrete(name="BMI Bin (kg/m^2)", labels=c("18.5-24", "25-29", "30>")) +
  scale_color_brewer(palette = "Dark2",name="Age Bin (yr)", labels=c("18-33", "34-49", "50-65")) +
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = 0, hjust = .25, vjust=0.5))
shannon_plot
```

```{r save shannon plot}
ggsave("../../plots/AUC_TMAO_Shannon_a_diversity_v3.jpg", shannon_plot,
       width = 8, 
       height = 5, 
       dpi = 300)
```


Differential abundance analysis via DESeq2

```{r DESeq2 by peakTMAO}
library(DESeq2)

# Get TMAO peak column and 
phenPeak <- phen[,"peakTMAO", F]

# Prune PSO to ones with TMAO peak data
keepSamp <- rownames(phenPeak)
PSOtmao1G_sub <- prune_samples(keepSamp, PSOtmao1G)
PSOtmao1G_sub

# Merge
phenPeak <- sample_data(phenPeak)
PSOpeak <- merge_phyloseq(PSOtmao1G_sub, phenPeak)
colnames(sample_data(PSOpeak))

# Set levels
sample_data(PSOpeak)$peakTMAO <- factor(sample_data(PSOpeak)$peakTMAO, levels = c("peak0", "peak30", "peak180", "peak360"))

# DESeq analysis - tmao_tertile
mb <- phyloseq_to_deseq2(PSOpeak, ~ sex + age + bmi_final.x + peakTMAO)
mb.ratio <- DESeq(mb, test = "LRT", reduced = ~ sex + age + bmi_final.x) # remove variable of interest here, TMAO
res.ratio <- results(mb.ratio, contrast = c("peakTMAO", "peak0", "peak360")) # Look at most different tertiles
alpha = 0.05
sigtab = res.ratio[which(res.ratio$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(PSOpeak)[rownames(sigtab), ], "matrix"))
sigtab$family_genus <- paste0(sigtab$Family, sigtab$Genus)
head(sigtab)

res.ratio2 <- results(mb.ratio, contrast = c("peakTMAO", "peak0", "peak30")) # Look at 0 and 180 grp
alpha = 0.05
sigtab2 = res.ratio2[which(res.ratio2$padj < alpha), ]
sigtab2 = cbind(as(sigtab2, "data.frame"), as(tax_table(PSOpeak)[rownames(sigtab2), ], "matrix"))
sigtab2$family_genus <- paste0(sigtab2$Family, sigtab2$Genus)
head(sigtab2)

```

```{r DESeq plot peakTMAO }
#~~~~~~~~~~~~~~~
# Plot Results
#~~~~~~~~~~~~~~~
resOTU <- rownames(sigtab)
otuData <- as.data.frame(otu_table(PSOpeak))
otuData_desRes <- otuData[rownames(otuData) %in% resOTU,]
#otuData_desRes <- as.data.frame(t(otuData_desRes))
# Get sample data groupings
sampleData <- as.data.frame(sample_data(PSOpeak))
sampleData <- subset(sampleData, select = c(peakTMAO, tmao_mdn, tmao_quantile, tmao_tertile, tmao))
# Get tax table
taxaData <- as.data.frame(tax_table(PSOpeak))
taxaData_res <- taxaData[rownames(taxaData) %in% resOTU,]
taxaData_res$FamilyGenus <- paste0(taxaData_res$Family, taxaData_res$Genus)
# Merge
ggData <- merge(taxaData_res, otuData_desRes, by = 0)
rownames(ggData) <-  ggData$FamilyGenus
ggData <- ggData[,-(1:9)]
ggData <- as.data.frame(t(ggData))
ggData <- merge(ggData, sampleData, by = 0)
rownames(ggData) <- ggData[,"Row.names"]
ggData <- subset(ggData, select = -c(Row.names))
# Save a copy of results for plotting
ggData_G <- ggData

# Loop Plot to view
nres <- length(resOTU)
comparisons <- list(c("peak0", "peak30"), c("peak0", "peak180"), c("peak0","peak360"),  c("peak30","peak180"),  c("peak30","peak360"),  c("peak180","peak360") )
for (i in 1:nres) { 
    print(ggplot(ggData_G, aes(x = peakTMAO, y = log(ggData_G[,i]))) +
            geom_boxplot(aes(color=peakTMAO)) +
            geom_point() +
            theme_light() +
            ylab(paste0("Log of", colnames(ggData_G[i]))) +
            xlab("TMAO Response Type") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "TMAO Response Type") +
            stat_compare_means(comparisons=comparisons)
    )
}


presentationPlot <- ggplot(ggData_G, aes(x = peakTMAO, y = log(ggData_G[,1]))) +
            geom_boxplot(aes(color=peakTMAO)) +
            geom_point() +
            theme_light() +
            ylab(paste0("Log of", colnames(ggData_G[1]))) +
            xlab("TMAO Response Type") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "TMAO Response Type")

presentationPlot

# Get OTU table info
taxaData <- as.data.frame(tax_table(PSOpeak))
unique(taxaData$Family)
unique(taxaData$Genus)
```

```{r save plot}
ggsave("../../plots/AUC_TMAO_DESeq_Prevotella.jpg", presentationPlot,
       width = 8, 
       height = 5, 
       dpi = 300)

```

Repeat looking at AUC-TMAO median
```{r DESeq2 by AUC median}
library(DESeq2)

# Make median variable
median(phen$AUC_TMAO_uM_hour)
phen$AUC_TMAO_mdn <- ifelse(phen$AUC_TMAO_uM_hour <= median(phen$AUC_TMAO_uM_hour), "less", "greater")
str(phen$AUC_TMAO_mdn)
phen$AUC_TMAO_mdn <- factor(phen$AUC_TMAO_mdn, levels = c("less", "greater"), labels = c("less", "greater"))

# Get TMAO peak column and 
phenPeak <- phen[,"AUC_TMAO_mdn", F]

# Prune PSO to ones with TMAO peak data
keepSamp <- rownames(phenPeak)
PSOtmao1G_sub.AUC <- prune_samples(keepSamp, PSOtmao1G)

# Merge
phenPeak <- sample_data(phenPeak)
PSO.AUC <- merge_phyloseq(PSOtmao1G_sub.AUC, phenPeak)
colnames(sample_data(PSO.AUC))

# Set levels
sample_data(PSO.AUC)$AUC_TMAO_mdn <- factor(sample_data(PSO.AUC)$AUC_TMAO_mdn, levels = c("less", "greater"))
sample_data(PSO.AUC)$screen_sex <- factor(sample_data(PSO.AUC)$screen_sex, levels = c("Female", "Male"))
check <- sample_data(PSO.AUC)
str(check$screen_sex) # male=2
str(check$AUC_TMAO_mdn)

# DESeq analysis - tmao_tertile
mb <- phyloseq_to_deseq2(PSO.AUC, ~ screen_sex + age + bmi_final.x + AUC_TMAO_mdn)
mb.ratio <- DESeq(mb, test = "LRT", reduced = ~ screen_sex + age + bmi_final.x) # remove variable of interest here, TMAO
res.ratio <- results(mb.ratio, contrast = c("AUC_TMAO_mdn", "less", "greater")) # Look at most different tertiles
alpha = 0.05
sigtab = res.ratio[which(res.ratio$padj < alpha), ]
sigtab = cbind(as(sigtab, "data.frame"), as(tax_table(PSO.AUC)[rownames(sigtab), ], "matrix"))
sigtab$family_genus <- paste0(sigtab$Family, sigtab$Genus)
head(sigtab)

```

Save
```{r}
write.csv(sigtab, "../../tables/Microbiome_AUCTMAO_DESeq2.csv")
```

```{r DESeq plot AUC-TMAO}
#~~~~~~~~~~~~~~~
# Plot Results
#~~~~~~~~~~~~~~~
resOTU <- rownames(sigtab)
otuData <- as.data.frame(otu_table(PSO.AUC))
otuData_desRes <- otuData[rownames(otuData) %in% resOTU,]
#otuData_desRes <- as.data.frame(t(otuData_desRes))
# Get sample data groupings
sampleData <- as.data.frame(sample_data(PSO.AUC))
sampleData <- subset(sampleData, select = c(AUC_TMAO_mdn))
# Get tax table
taxaData <- as.data.frame(tax_table(PSO.AUC))
taxaData_res <- taxaData[rownames(taxaData) %in% resOTU,]
taxaData_res$FamilyGenus <- paste0(taxaData_res$Family, taxaData_res$Genus)
# Merge
ggData <- merge(taxaData_res, otuData_desRes, by = 0)
rownames(ggData) <-  ggData$FamilyGenus
ggData <- ggData[,-(1:9)]
ggData <- as.data.frame(t(ggData))
ggData <- merge(ggData, sampleData, by = 0)
rownames(ggData) <- ggData[,"Row.names"]
ggData <- subset(ggData, select = -c(Row.names))
# Save a copy of results for plotting
ggData_G <- ggData

# Loop Plot to view
nres <- length(resOTU)
comparisons <- list(c("less", "greater") )
for (i in 1:nres) { 
    print(ggplot(ggData_G, aes(x = AUC_TMAO_mdn, y = log(ggData_G[,i]))) +
            geom_boxplot(aes(color=AUC_TMAO_mdn)) +
            geom_point() +
            theme_bw() +
            ylab(paste0("Log of", colnames(ggData_G[i]))) +
            xlab("AUC-TMAO Median Group") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "TMAO Response Type") +
            stat_compare_means(comparisons=comparisons, method = "wilcox.test")
    )
}


presentationPlot <- ggplot(ggData_G, aes(x = AUC_TMAO_mdn, y = log(ggData_G[,1]))) +
            geom_boxplot(aes(color=AUC_TMAO_mdn)) +
            geom_point() +
            theme_bw() +
            ylab(paste0("Log of", colnames(ggData_G[1]))) +
            xlab("AUC-TMAO Median Group") +
            scale_color_brewer(palette = "Dark2") +
            labs(color = "AUC-TMAO Median Group")

presentationPlot

# All results plot
sigtabgen = subset(sigtab, !is.na(Genus))
# Phylum order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Phylum = factor(as.character(sigtabgen$Phylum), levels=names(x))
# Genus order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Genus = factor(as.character(sigtabgen$Genus), levels=names(x))
DESeqPlot <- ggplot(sigtabgen, aes(y=family_genus, x=log2FoldChange, color=Phylum)) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3) + 
  ylab("Taxonomy") +
  xlab("Log2 Fold Change") + 
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = 0),
        axis.text.y = element_text(angle = 0))
DESeqPlot

# Save
tiff("../../plots/DESeq2_Genus_AUCTMAOmedian_221031.tiff", width = 8, height = 4, units = 'in', res = 300)
DESeqPlot
dev.off()

# How many non-0s?
sum(ggData_G$` f__Prevotellaceae g__Prevotella` != 0) #46
sum(ggData_G$` f__Veillonellaceae g__Dialister` != 0) # 37
sum(ggData_G$` f__Coriobacteriaceae g__Eggerthella` != 0) # 23
sum(ggData_G$` f__Erysipelotrichaceae g__Catenibacterium` != 0) # 19
sum(ggData_G$` f__Erysipelotrichaceae g__Clostridium` != 0) # 57
sum(ggData_G$` f__Lactobacillaceae g__Lactobacillus` != 0) # 19

# % range in cohort 
(19/97)*100 #19.6%
(57/97)*100 #58.8%
```

```{r beta diversity}
library(vegan)

# PCoA
PCoA_WU <- ordinate(PSOpeak, "PCoA", "wunifrac")
PCoA_UnWU <- ordinate(PSOpeak, "PCoA", "uunifrac")
PCoA_BC <- ordinate(PSOpeak, "PCoA", "bray")

# Calculate distance matrix
dismax_wu = phyloseq::distance(PSOpeak, method="wunifrac", type = "samples")
dismax_uwu = phyloseq::distance(PSOpeak, method="unifrac", type = "samples")
dismax_bc = phyloseq::distance(PSOpeak, method="bray", type = "samples")

# Beta dispersion
# Check AUC and peakTMAO variables
df.auc <- as(sample_data(PSO.AUC), "data.frame")
# WU
bdis <- betadisper(dismax_wu, df2$peakTMAO)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)
bdis <- betadisper(dismax_wu, df.auc$AUC_TMAO_mdn)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)
# UWU
bdis <- betadisper(dismax_uwu, df2$peakTMAO)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)
bdis <- betadisper(dismax_uwu, df.auc$AUC_TMAO_mdn)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)
# BC
bdis <- betadisper(dismax_bc, df2$peakTMAO)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)
bdis <- betadisper(dismax_bc, df.auc$AUC_TMAO_mdn)
permutest(bdis) # NS, good
boxplot(bdis, las = 2)

# ADONIS (perMANOVA) analysis
adonis2(dismax_wu ~ tmao_log + sex + age + bmi_final.y, data=df2, method = "wunifrac", by = "margin", perm=999) # df
adonis2(dismax_uwu ~ tmao_log + sex + age + bmi_final.y, data=df2, method = "unifrac", by = "margin", perm=999) # df
adonis2(dismax_bc ~ tmao_log + sex + age + bmi_final.y, data=df2, method = "bray", by = "margin", perm=999) # df
# note, the by = "margin" command makes the order of the covariates not matter in the model

# ADONIS (perMANOVA) analysis
adonis2(dismax_wu ~ log(AUC_TMAO_uM_hour) + sex + age + bmi_final.y, data=df2, method = "wunifrac", by = "margin", perm=999) # df
adonis2(dismax_uwu ~ log(AUC_TMAO_uM_hour) + sex + age + bmi_final.y, data=df2, method = "unifrac", by = "margin", perm=999) # df
adonis2(dismax_bc ~ log(AUC_TMAO_uM_hour) + sex + age + bmi_final.y, data=df2, method = "bray", by = "margin", perm=999) # df

# ADONIS (perMANOVA) analysis
adonis2(dismax_wu ~ peakTMAO + sex + age + bmi_final.y, data=df2, method = "wunifrac", by = "margin", perm=999) # df
adonis2(dismax_uwu ~ peakTMAO + sex + age + bmi_final.y, data=df2, method = "unifrac", by = "margin", perm=999) # df
adonis2(dismax_bc ~ peakTMAO + sex + age + bmi_final.y, data=df2, method = "bray", by = "margin", perm=999) # df

# Weighted
betaplot_WU <- plot_ordination(PSOpeak, PCoA_WU, 
                    color="peakTMAO",
                    shape = NULL,
                    axes = 1:2) +
  scale_color_brewer(palette = "Dark2", labels = c("0", "30", "180", "360")) +
  stat_ellipse(linetype = 1) +
  labs(color = "Peak TMAO Group") +
  theme_bw() + 
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = 0, hjust = 0, vjust=0))

  betaplot_WU

ggsave("../../plots/PeakTMAO_WU_b_diversity.jpg", betaplot_WU,
       width = 8, 
       height = 5, 
       dpi = 300)
  
# UnWeighted
betaplot_UnWU <- plot_ordination(PSOpeak, PCoA_UnWU, 
                    color="peakTMAO",
                    shape = NULL,
                    axes = 1:2) +
  scale_color_brewer(palette = "Dark2", labels = c("0", "30", "180", "360")) +
  stat_ellipse(linetype = 1) +
  labs(color = "Peak TMAO Group") +
  theme_bw() + 
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

betaplot_UnWU

ggsave("../../plots/PeakTMAO_UnWU_b_diversity.jpg", betaplot_UnWU,
       width = 8, 
       height = 5, 
       dpi = 300)

# BC
betaplot_BC <- plot_ordination(PSOpeak, PCoA_BC, 
                    color="peakTMAO",
                    shape = NULL,
                    axes = 1:2) +
  scale_color_brewer(palette = "Dark2", labels = c("0", "30", "180", "360")) +
  stat_ellipse(linetype = 1) +
  labs(color = "Peak TMAO Group") +
  theme_minimal() + 
  theme(axis.text = element_text(size = 12), # x and y axis text size
        axis.title = element_text(size = 12), # x and y label text size
        legend.text = element_text(size = 12), # legend text size
        legend.title = element_text(size = 12), # label legend text size
        axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))

betaplot_BC

# NMDS
PSOpeak.ord <- ordinate(PSOpeak, "NMDS", "bray")
p1 = plot_ordination(PSOpeak, PSOpeak.ord, type="samples", color="peakTMAO", title="samples") + 
  stat_ellipse(linetype = 1) 
print(p1)
```

Looks a bit of a dead end. PeakTMAO group doesn't clearly differentiate microbiome composition.

Check Firm:Bact ratio

```{r Firmicutes to bacteroidetes ratio per peakTMAO group}
# Combine to Phylum level
PSOpeak_Phylum <- tax_glom(PSOpeak, "Phylum", NArm = TRUE)

# Subset by peak group
PSOpeak_Phylum_peak0 <- subset_samples(PSOpeak_Phylum, peakTMAO == "peak0")
PSOpeak_Phylum_peak30 <- subset_samples(PSOpeak_Phylum, peakTMAO == "peak30")
PSOpeak_Phylum_peak180 <- subset_samples(PSOpeak_Phylum, peakTMAO == "peak180")
PSOpeak_Phylum_peak360 <- subset_samples(PSOpeak_Phylum, peakTMAO == "peak360")

# View
PSOpeak_Phylum_peak0
PSOpeak_Phylum_peak30
PSOpeak_Phylum_peak180
PSOpeak_Phylum_peak360

#~~~~~~~~~~
# Peak0
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalence_Phy_peak0 = apply(X = otu_table(PSOpeak_Phylum_peak0),
               MARGIN = ifelse(taxa_are_rows(PSOpeak_Phylum_peak0), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
prevalence_Phy_peak0 = data.frame(Prevalence = prevalence_Phy_peak0,
                    TotalAbundance = taxa_sums(PSOpeak_Phylum_peak0), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOpeak_Phylum_peak0))
# Firmicutes:Bacteroidetes Abundance Ratio
# Note, calculating it this way takes total abundance of firm and total abundance of bact in the peak0 group and makes a ratio of them. It doesn't take the mean of every person's ratio.  
firm_bacter_peak0 <- prevalence_Phy_peak0$TotalAbundance[prevalence_Phy_peak0$Phylum == " p__Firmicutes"] / prevalence_Phy_peak0$TotalAbundance[prevalence_Phy_peak0$Phylum == " p__Bacteroidetes"] 

# To get a mean of every person's ratio: 
# This method also lets you get statistics like SD
# Melt PSO data
mlt_PSOpeak_Phylum_peak0 <- psmelt(PSOpeak_Phylum_peak0)
# Get and format tax table
tbl.mlt_PSOpeak_Phylum_peak0 <- as.data.frame(tax_table(PSOpeak_Phylum_peak0))
tbl.mlt_PSOpeak_Phylum_peak0 <-  tbl.mlt_PSOpeak_Phylum_peak0[,1:2]
tbl.mlt_PSOpeak_Phylum_peak0$OTU <- row.names(tbl.mlt_PSOpeak_Phylum_peak0)
# Merge
all.mlt_PSOpeak_Phylum_peak0 <- merge(tbl.mlt_PSOpeak_Phylum_peak0, mlt_PSOpeak_Phylum_peak0, by = "OTU", all.y = TRUE)
# Manipulate to get stats
# only keep firm and bact
firm_peak0 <- all.mlt_PSOpeak_Phylum_peak0[all.mlt_PSOpeak_Phylum_peak0$Phylum.x == " p__Firmicutes",]
firm_peak0 <- firm_peak0[,c("Sample", "Abundance")]
colnames(firm_peak0) <- paste0(colnames(firm_peak0), "_firm")
bact_peak0 <- all.mlt_PSOpeak_Phylum_peak0[all.mlt_PSOpeak_Phylum_peak0$Phylum.x == " p__Bacteroidetes",]
bact_peak0 <- bact_peak0[,c("Sample", "Abundance")]
colnames(bact_peak0) <- paste0(colnames(bact_peak0), "_bact")
# Merge
fb_peak0 <- merge(firm_peak0, bact_peak0, by.x = "Sample_firm", by.y = "Sample_bact")
fb_peak0$FBratio <- (fb_peak0$Abundance_firm / fb_peak0$Abundance_bact)
# Mean and sd
mean(fb_peak0$FBratio) #8.25
sd(fb_peak0$FBratio) #11.44

#~~~~~~~~~~
# Peak 30
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalence_Phy_peak30 = apply(X = otu_table(PSOpeak_Phylum_peak30),
               MARGIN = ifelse(taxa_are_rows(PSOpeak_Phylum_peak30), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
prevalence_Phy_peak30 = data.frame(Prevalence = prevalence_Phy_peak30,
                    TotalAbundance = taxa_sums(PSOpeak_Phylum_peak30), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOpeak_Phylum_peak30))
firm_bacter_peak30 <- prevalence_Phy_peak30$TotalAbundance[prevalence_Phy_peak30$Phylum == " p__Firmicutes"] / prevalence_Phy_peak30$TotalAbundance[prevalence_Phy_peak30$Phylum == " p__Bacteroidetes"] 

# To get a mean of every person's ratio: 
# This method also lets you get statistics like SD
# Melt PSO data
mlt_PSOpeak_Phylum_peak30 <- psmelt(PSOpeak_Phylum_peak30)
# Get and format tax table
tbl.mlt_PSOpeak_Phylum_peak30 <- as.data.frame(tax_table(PSOpeak_Phylum_peak30))
tbl.mlt_PSOpeak_Phylum_peak30 <-  tbl.mlt_PSOpeak_Phylum_peak30[,1:2]
tbl.mlt_PSOpeak_Phylum_peak30$OTU <- row.names(tbl.mlt_PSOpeak_Phylum_peak30)
# Merge
all.mlt_PSOpeak_Phylum_peak30 <- merge(tbl.mlt_PSOpeak_Phylum_peak30, mlt_PSOpeak_Phylum_peak30, by = "OTU", all.y = TRUE)
# Manipulate to get stats
# only keep firm and bact
firm_peak30 <- all.mlt_PSOpeak_Phylum_peak30[all.mlt_PSOpeak_Phylum_peak30$Phylum.x == " p__Firmicutes",]
firm_peak30 <- firm_peak30[,c("Sample", "Abundance")]
colnames(firm_peak30) <- paste0(colnames(firm_peak30), "_firm")
bact_peak30 <- all.mlt_PSOpeak_Phylum_peak30[all.mlt_PSOpeak_Phylum_peak30$Phylum.x == " p__Bacteroidetes",]
bact_peak30 <- bact_peak30[,c("Sample", "Abundance")]
colnames(bact_peak30) <- paste0(colnames(bact_peak30), "_bact")
# Merge
fb_peak30 <- merge(firm_peak30, bact_peak30, by.x = "Sample_firm", by.y = "Sample_bact")
fb_peak30$FBratio <- (fb_peak30$Abundance_firm / fb_peak30$Abundance_bact)
# Mean and sd
mean(fb_peak30$FBratio) #10.28
sd(fb_peak30$FBratio) #12.94


#~~~~~~~~~~
# Peak 180
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalence_Phy_peak180 = apply(X = otu_table(PSOpeak_Phylum_peak180),
               MARGIN = ifelse(taxa_are_rows(PSOpeak_Phylum_peak180), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
prevalence_Phy_peak180 = data.frame(Prevalence = prevalence_Phy_peak180,
                    TotalAbundance = taxa_sums(PSOpeak_Phylum_peak180), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOpeak_Phylum_peak180))
# Firmicutes:Bacteroidetes Abundance Ratio
firm_bacter_peak180 <- prevalence_Phy_peak180$TotalAbundance[prevalence_Phy_peak180$Phylum == " p__Firmicutes"] / prevalence_Phy_peak180$TotalAbundance[prevalence_Phy_peak180$Phylum == " p__Bacteroidetes"] 

# To get a mean of every person's ratio: 
# This method also lets you get statistics like SD
# Melt PSO data
mlt_PSOpeak_Phylum_peak180 <- psmelt(PSOpeak_Phylum_peak180)
# Get and format tax table
tbl.mlt_PSOpeak_Phylum_peak180 <- as.data.frame(tax_table(PSOpeak_Phylum_peak180))
tbl.mlt_PSOpeak_Phylum_peak180 <-  tbl.mlt_PSOpeak_Phylum_peak180[,1:2]
tbl.mlt_PSOpeak_Phylum_peak180$OTU <- row.names(tbl.mlt_PSOpeak_Phylum_peak180)
# Merge
all.mlt_PSOpeak_Phylum_peak180 <- merge(tbl.mlt_PSOpeak_Phylum_peak180, mlt_PSOpeak_Phylum_peak180, by = "OTU", all.y = TRUE)
# Manipulate to get stats
# only keep firm and bact
firm_peak180 <- all.mlt_PSOpeak_Phylum_peak180[all.mlt_PSOpeak_Phylum_peak180$Phylum.x == " p__Firmicutes",]
firm_peak180 <- firm_peak180[,c("Sample", "Abundance")]
colnames(firm_peak180) <- paste0(colnames(firm_peak180), "_firm")
bact_peak180 <- all.mlt_PSOpeak_Phylum_peak180[all.mlt_PSOpeak_Phylum_peak180$Phylum.x == " p__Bacteroidetes",]
bact_peak180 <- bact_peak180[,c("Sample", "Abundance")]
colnames(bact_peak180) <- paste0(colnames(bact_peak180), "_bact")
# Merge
fb_peak180 <- merge(firm_peak180, bact_peak180, by.x = "Sample_firm", by.y = "Sample_bact")
fb_peak180$FBratio <- (fb_peak180$Abundance_firm / fb_peak180$Abundance_bact)
# Mean and sd
mean(fb_peak180$FBratio) #8.82
sd(fb_peak180$FBratio) #8.94

#~~~~~~~~~~
# Peak 360
#~~~~~~~~~~
# Prevalence - how many samples the taxa is observed in
prevalence_Phy_peak360 = apply(X = otu_table(PSOpeak_Phylum_peak360),
               MARGIN = ifelse(taxa_are_rows(PSOpeak_Phylum_peak360), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
prevalence_Phy_peak360 = data.frame(Prevalence = prevalence_Phy_peak360,
                    TotalAbundance = taxa_sums(PSOpeak_Phylum_peak360), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOpeak_Phylum_peak360))
firm_bacter_peak360 <- prevalence_Phy_peak360$TotalAbundance[prevalence_Phy_peak360$Phylum == " p__Firmicutes"] / prevalence_Phy_peak360$TotalAbundance[prevalence_Phy_peak360$Phylum == " p__Bacteroidetes"] 

# To get a mean of every person's ratio: 
# This method also lets you get statistics like SD
# Melt PSO data
mlt_PSOpeak_Phylum_peak360 <- psmelt(PSOpeak_Phylum_peak360)
# Get and format tax table
tbl.mlt_PSOpeak_Phylum_peak360 <- as.data.frame(tax_table(PSOpeak_Phylum_peak360))
tbl.mlt_PSOpeak_Phylum_peak360 <-  tbl.mlt_PSOpeak_Phylum_peak360[,1:2]
tbl.mlt_PSOpeak_Phylum_peak360$OTU <- row.names(tbl.mlt_PSOpeak_Phylum_peak360)
# Merge
all.mlt_PSOpeak_Phylum_peak360 <- merge(tbl.mlt_PSOpeak_Phylum_peak360, mlt_PSOpeak_Phylum_peak360, by = "OTU", all.y = TRUE)
# Manipulate to get stats
# only keep firm and bact
firm_peak360 <- all.mlt_PSOpeak_Phylum_peak360[all.mlt_PSOpeak_Phylum_peak360$Phylum.x == " p__Firmicutes",]
firm_peak360 <- firm_peak360[,c("Sample", "Abundance")]
colnames(firm_peak360) <- paste0(colnames(firm_peak360), "_firm")
bact_peak360 <- all.mlt_PSOpeak_Phylum_peak360[all.mlt_PSOpeak_Phylum_peak360$Phylum.x == " p__Bacteroidetes",]
bact_peak360 <- bact_peak360[,c("Sample", "Abundance")]
colnames(bact_peak360) <- paste0(colnames(bact_peak360), "_bact")
# Merge
fb_peak360 <- merge(firm_peak360, bact_peak360, by.x = "Sample_firm", by.y = "Sample_bact")
fb_peak360$FBratio <- (fb_peak360$Abundance_firm / fb_peak360$Abundance_bact)
# Mean and sd
mean(fb_peak360$FBratio) #6.70
sd(fb_peak360$FBratio) #8.19

#~~~~~~~~~~
# Are means of ratios from individual people in group statistically different?
#~~~~~~~~~~
fb_peak0$PeakGrp <- "peak0"
fb_peak30$PeakGrp <- "peak30"
fb_peak180$PeakGrp <- "peak180"
fb_peak360$PeakGrp <- "peak360"
# rbind
fb_all <- rbind(fb_peak0,fb_peak30,fb_peak180,fb_peak360)
summary(aov(FBratio ~ PeakGrp, fb_all))

#~~~~~~~~~~
# Compare ratios
#~~~~~~~~~~
# from total abundance of group
cat("Firmicutes:Bacteroidetes ratio peak0:", firm_bacter_peak0)
cat("Firmicutes:Bacteroidetes ratio peak30:", firm_bacter_peak30)
cat("Firmicutes:Bacteroidetes ratio peak180:", firm_bacter_peak180)
cat("Firmicutes:Bacteroidetes ratio peak360:", firm_bacter_peak360)
# mean of ratios from individual people in group 
cat("Firmicutes:Bacteroidetes ratio peak0 from person means:", mean(fb_peak0$FBratio))
cat("Firmicutes:Bacteroidetes ratio peak30 from person means:", mean(fb_peak30$FBratio))
cat("Firmicutes:Bacteroidetes ratio peak180 from person means:", mean(fb_peak180$FBratio))
cat("Firmicutes:Bacteroidetes ratio peak360 from person means:", mean(fb_peak360$FBratio))
```

Are firmicutes related to AUC-TMAO?
```{r firmicutes and AUC-TMAO}
# Melt PSO data
mlt_PSOpeak_Phylum <- psmelt(PSOpeak_Phylum)
# Get and format tax table
tbl.mlt_PSOpeak_Phylum <- as.data.frame(tax_table(PSOpeak_Phylum))
tbl.mlt_PSOpeak_Phylum <-  tbl.mlt_PSOpeak_Phylum[,1:2]
tbl.mlt_PSOpeak_Phylum$OTU <- row.names(tbl.mlt_PSOpeak_Phylum)
# Merge
all.mlt_PSOpeak_Phylum <- merge(tbl.mlt_PSOpeak_Phylum, mlt_PSOpeak_Phylum, by = "OTU", all.y = TRUE)
# Keep firm
firm <- all.mlt_PSOpeak_Phylum[all.mlt_PSOpeak_Phylum$Phylum.x == " p__Firmicutes",]
rownames(firm) <- firm$Sample

# Merge with AUC-TMAO
firm.df2 <- merge(df2, firm, by = 0)

# stats
colnames(firm.df2)
shapiro.test(firm.df2$Abundance)
shapiro.test(log(firm.df2$Abundance))
summary(lm(log(AUC_TMAO_uM_hour) ~ log(Abundance), firm.df2))

```


Plot family level relative abundance
```{r prevalence plot}
# Prevalence - how many samples the taxa is observed in
prevdf = apply(X = otu_table(PSOpeak),
               MARGIN = ifelse(taxa_are_rows(PSOpeak), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})
prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(PSOpeak), # the total number of occurrences (even if only observed 50 times in 1 sample, the number would be 50)
                    tax_table(PSOpeak))
print("Table showing prevalence, total abundance, and taxonomy:")
prevdf[1:10,]


# Family
# Agglomerate taxa
PSOpeak_F = tax_glom(PSOpeak, "Family", NArm = TRUE)
PSOpeak_F <- microbiome::aggregate_top_taxa(PSOpeak_F, "Family", top = 10)
ps1F.10.family.comp <- microbiome::transform(PSOpeak_F, transform="compositional") #compositional

# Plot
plot.composition.relAbun <- microbiome::plot_composition(ps1F.10.family.comp,
                                             sample.sort = "Description",
                                             #x.label = "peakTMAO",
                                             group_by = "peakTMAO",
                                             average_by = "peakTMAO") +
  theme_bw() +
  upstartr::scale_y_percent() + 
  scale_fill_brewer("Family", palette = "Paired") + 
  scale_x_discrete(breaks=c("peak0", "peak30", "peak180", "peak360"),
        labels=c("Peak 0m", "Peak 30m", "Peak 3hr", "Peak 6hr")) + 
  labs(x= "Peak TMAO Group", y = "Relative Abundance (%)")

plot.composition.relAbun
# Save
tiff("../../plots/RelAbund_Fam_peakTMAO_221031.tiff", width = 5, height = 4, units = 'in', res = 300)
plot.composition.relAbun
dev.off()
```



