---
title: "Phosphatidylcholine Exploration"
author: "Kristen James"
date: "2022-09-13"
output: html_document
---

Could the phosphatidylcholines be providing a TMAO-precursor in response to digestion?

9/13/22 - One way to address this, ask if the metabolites are differentially present by peak-TMAO group at each blood draw

10/11/22 - Lets look at the sum of the phosphatidylcholines. Do those differ by peak-TMAO group?

```{r libraries}
library(dplyr)
```


```{r}
# Clear environment to start fresh (if necessary)
rm(list = ls())

# Data from TMAO_Meal_Response_Focused.rmd
#dfwide <- readRDS("../../Data/Processed/dfwide_metavariables_n97_inflam.rds")

# Long data
dflong <- readRDS("../../Data/Processed/Biocrates_long_meta.rds")

dflong$pp_Time_min <- factor(dflong$pp_Time_min, levels = c("0", "30", "180", "360"))

#BMI
dflong$bmi_bin <- dflong$bin_number
dflong$bmi_bin <- gsub("..","", dflong$bmi_bin)
#Age
dflong$age_bin <- dflong$bin_number
dflong$age_bin <- gsub("^.","", dflong$age_bin)
dflong$age_bin <- gsub(".$","", dflong$age_bin)
#Sex
dflong$sex_bin <- dflong$bin_number
dflong$sex_bin <- gsub("..$","", dflong$sex_bin)

# Put age and sex at beginning
dflong <- dflong %>% select(c(age, sex), everything())
colnames(dflong)
# Cut non-metabolite out from end
dflong <- dflong[,1:593]
# Make sex 0 or 1
dflong$sex <- ifelse(dflong$sex == 1, 0, 1) # before 1=male,2=female. After 0=male, 1=female
table(dflong$sex)

# Wide data
dfwide <- readRDS("../../Data/Processed/Biocrates_AreaUnderCurve_meta.rds")
rownames(dfwide) <- dfwide[,"Subject_ID"]
# peak TMAO classification
dfwide$peakTMAO <- ifelse(dfwide$TMAO.0 >= dfwide$TMAO.30 & 
                            dfwide$TMAO.0 >= dfwide$TMAO.180 & 
                              dfwide$TMAO.0 >= dfwide$TMAO.360, "peak0",
                          ifelse(dfwide$TMAO.30 >= dfwide$TMAO.0 &
                                   dfwide$TMAO.30 >= dfwide$TMAO.180 &
                                    dfwide$TMAO.30 >= dfwide$TMAO.360, "peak30", 
                                 ifelse(dfwide$TMAO.180 >= dfwide$TMAO.0 &
                                          dfwide$TMAO.180 >= dfwide$TMAO.30 &
                                            dfwide$TMAO.180 >= dfwide$TMAO.360, "peak180", "peak360")))

# Set levels
dfwide$peakTMAO <- factor(dfwide$peakTMAO, levels = c("peak0", "peak30", "peak180", "peak360"))
# Distribution of ppl by peak TMAO concentration
table(dfwide$peakTMAO)

# get list of ids by peakTMAO group
peakTMAO <- dfwide[,c("peakTMAO"),drop=FALSE]

# Add list to dflong
dflong <- merge(peakTMAO, dflong, by.x = 0, by.y = "Subject_ID", all.y = TRUE)

# Only keep IDs in peakTMAO
dflongPk <- dflong[dflong$Row.names %in% rownames(peakTMAO),]

colnames(dflongPk)

# Make blood draw numeric and showing the right relationships with regard to time
dflongPk2 <- dflongPk
dflongPk2$time_numeric <- ifelse(dflongPk2$Blood_Draw == "BD1", 0,
                              ifelse(dflongPk2$Blood_Draw == "BD2", 30,
                                     ifelse(dflongPk2$Blood_Draw == "BD3", 180, 360)))
```

Is the metabolite different by peakTMAO group?
- Get a blood draw
- Get a metabolite
- Run the stats
- Organize the answer
- Repeat

Note the chunk below in *interactive*. You can't just hit the green arrow!

```{r}
# Name blood draw
# BD1
bd1 <- dflongPk[dflongPk$Blood_Draw == "BD1",]
bd2 <- dflongPk[dflongPk$Blood_Draw == "BD2",]
bd3 <- dflongPk[dflongPk$Blood_Draw == "BD3",]
bd4 <- dflongPk[dflongPk$Blood_Draw == "BD4",]

# Set limits
n_metabolites <- ncol(dflongPk[,12:594])

# Now run... some steps need to be changed to specify bd. Look for comments below indicating places to change accordingly
# Change here with BD
metabolite <- bd1[,12:594]
metabolite <- bd2[,12:594]
metabolite <- bd3[,12:594]
metabolite <- bd4[,12:594]

# Check
str(dflongPk$sex)
str(dflongPk$age)
str(dflongPk$peakTMAO)

# Linear model
my_lm <- lapply(1:n_metabolites, function(x) lm(metabolite[,x] ~ peakTMAO + sex*age, bd2 )) # change bd with correct iteration

# Get statistics
my_lm_adj.r.squared <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm_estimate <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm_std.error <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm_T.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm_P.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
  
# Organize results
my_lm_temp <- cbind(my_lm_adj.r.squared, my_lm_estimate, my_lm_std.error, my_lm_T.value, my_lm_P.value)
# Set colnames and rownames
colnames(my_lm_temp) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm_temp) <- colnames(metabolite)
# Data frame
res_metabolite <- as.data.frame(my_lm_temp)

# BH correction
res_metabolite$p.adj <- stats::p.adjust(res_metabolite$P, method = "BH", n = n_metabolites)

# Save
res_metabolite_bd1 <- res_metabolite
res_metabolite_bd2 <- res_metabolite
res_metabolite_bd3 <- res_metabolite
res_metabolite_bd4 <- res_metabolite
```

We are most interested in PC, LysPC, BA. 
Note the chunk below in *interactive*. You can't just hit the green arrow!

```{r}
# Select bile acids and phosphatidylcholines
colnames(dflongPk)
# bileacids = 91-104
# phosphatidylchohlines = 216-303
# TMAO = 49
# peakTMAO = 2
# sex = 3
# age = 4
# blood draw = 9
dflongPk_sub <- dflongPk[,c(2:4,9,49,91:104,216:303)]

sub_bd1 <- dflongPk_sub[dflongPk_sub$Blood_Draw == "BD1",]
sub_bd2 <- dflongPk_sub[dflongPk_sub$Blood_Draw == "BD2",]
sub_bd3 <- dflongPk_sub[dflongPk_sub$Blood_Draw == "BD3",]
sub_bd4 <- dflongPk_sub[dflongPk_sub$Blood_Draw == "BD4",]

# Set limits
colnames(dflongPk_sub)
n_metabolites <- ncol(dflongPk_sub[,5:107])

# Now run... some steps need to be changed to specify bd. Look for comments below indicating places to change accordingly
# Change here with BD
colnames(sub_bd1)
metabolite <- sub_bd1[,5:107]
metabolite <- sub_bd2[,5:107]
metabolite <- sub_bd3[,5:107]
metabolite <- sub_bd4[,5:107]

# Check
str(sub_bd1$sex)

# Linear model
my_lm <- lapply(1:n_metabolites, function(x) lm(metabolite[,x] ~ peakTMAO + sex*age, sub_bd4 )) # change bd with correct iteration

# Get statistics
my_lm_adj.r.squared <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm_estimate <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm_std.error <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm_T.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm_P.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
  
# Organize results
my_lm_temp <- cbind(my_lm_adj.r.squared, my_lm_estimate, my_lm_std.error, my_lm_T.value, my_lm_P.value)
# Set colnames and rownames
colnames(my_lm_temp) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm_temp) <- colnames(metabolite)
# Data frame
res_metabolite <- as.data.frame(my_lm_temp)

# BH correction
res_metabolite$p.adj <- stats::p.adjust(res_metabolite$P, method = "BH", n = n_metabolites)

# Save
res_metabolite_bd1_sub <- res_metabolite
res_metabolite_bd2_sub <- res_metabolite
res_metabolite_bd3_sub <- res_metabolite
res_metabolite_bd4_sub <- res_metabolite

res_metabolite_bd1_sub 
res_metabolite_bd2_sub 
res_metabolite_bd3_sub 
res_metabolite_bd4_sub 
# Bile acids are different by peak-TMAO group and by blood draw
```

Bile Acid Exploration:

What is the relationship between bile acids and peak TMAO group? Correlation may be okay

```{r}
# Need AUC_TMAO_uM_hour combined with bile acid data by blood draw
AUC_TMAO <- dfwide[,"AUC_TMAO_uM_hour",drop=FALSE]

# Get BA df (all BDs)
dflongPk_BA <- dflongPk[,c(1,3:4,9,91:104)]

# merge AUC
dflongPk_BA_AUC <- merge(AUC_TMAO, dflongPk_BA, by.x = 0, by.y = "Row.names", all.y = TRUE)

# Still have to take a BD by BD approach
sub_bd1 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD1",]
sub_bd2 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD2",]
sub_bd3 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD3",]
sub_bd4 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD4",]

# BD1
colnames(sub_bd1)
b=apply(sub_bd1[,c(2,6:19)],2,function(x){
  cor.test(sub_bd1[,"AUC_TMAO_uM_hour"], x, method="spearman", adjust="BH")
})
p.val <- sapply(b,"[[","p.value")
head(p.val)  
r.cor <- sapply(b,"[[","estimate")
head(r.cor)
p <- cbind(p.val)
r <- cbind(r.cor)
pr <- cbind(p,r)
pr <- as.data.frame(pr)
AUC_cor_bd1 <- pr
AUC_cor_bd1

# BD2
colnames(sub_bd2)
b=apply(sub_bd2[,c(2,6:19)],2,function(x){
  cor.test(sub_bd2[,"AUC_TMAO_uM_hour"], x, method="spearman", adjust="BH")
})
p.val <- sapply(b,"[[","p.value")
head(p.val)  
r.cor <- sapply(b,"[[","estimate")
head(r.cor)
p <- cbind(p.val)
r <- cbind(r.cor)
pr <- cbind(p,r)
pr <- as.data.frame(pr)
AUC_cor_bd2 <- pr
AUC_cor_bd2

# BD3
colnames(sub_bd3)
b=apply(sub_bd3[,c(2,6:19)],2,function(x){
  cor.test(sub_bd3[,"AUC_TMAO_uM_hour"], x, method="spearman", adjust="BH")
})
p.val <- sapply(b,"[[","p.value")
head(p.val)  
r.cor <- sapply(b,"[[","estimate")
head(r.cor)
p <- cbind(p.val)
r <- cbind(r.cor)
pr <- cbind(p,r)
pr <- as.data.frame(pr)
AUC_cor_bd3 <- pr
AUC_cor_bd3

# BD4
colnames(sub_bd4)
b=apply(sub_bd4[,c(2,6:19)],2,function(x){
  cor.test(sub_bd4[,"AUC_TMAO_uM_hour"], x, method="spearman", adjust="BH")
})
p.val <- sapply(b,"[[","p.value")
head(p.val)  
r.cor <- sapply(b,"[[","estimate")
head(r.cor)
p <- cbind(p.val)
r <- cbind(r.cor)
pr <- cbind(p,r)
pr <- as.data.frame(pr)
AUC_cor_bd4 <- pr
AUC_cor_bd4
```
Bile acids do not correlate well to AUC-TMAO in the full cohort.

Run lm adjusting for covariates.

```{r}
# Need AUC_TMAO_uM_hour combined with bile acid data by blood draw
AUC_TMAO <- dfwide[,"AUC_TMAO_uM_hour",drop=FALSE]

# Get BA df (all BDs)
dflongPk_BA <- dflongPk[,c(1,3:4,9,91:104)]

# merge AUC
dflongPk_BA_AUC <- merge(AUC_TMAO, dflongPk_BA, by.x = 0, by.y = "Row.names", all.y = TRUE)

# Still have to take a BD by BD approach
sub_bd1 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD1",]
sub_bd2 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD2",]
sub_bd3 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD3",]
sub_bd4 <- dflongPk_BA_AUC[dflongPk_BA_AUC$Blood_Draw == "BD4",]

# Set limits
colnames(dflongPk_BA_AUC)
n_metabolites <- ncol(dflongPk_BA_AUC[,6:19])

# Now run... some steps need to be changed to specify bd. Look for comments below indicating places to change accordingly
# Change here with BD
metabolite <- sub_bd1[,6:19]
my_lm <- lapply(1:n_metabolites, function(x) lm(metabolite[,x] ~ log(AUC_TMAO_uM_hour) + sex*age, sub_bd1 )) # change bd with correct iteration

# Get statistics
my_lm_adj.r.squared <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm_estimate <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm_std.error <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm_T.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm_P.value <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
  
# Organize results
my_lm_temp <- cbind(my_lm_adj.r.squared, my_lm_estimate, my_lm_std.error, my_lm_T.value, my_lm_P.value)
# Set colnames and rownames
colnames(my_lm_temp) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm_temp) <- colnames(metabolite)
# Data frame
res_metabolite <- as.data.frame(my_lm_temp)

# BH correction
res_metabolite$p.adj <- stats::p.adjust(res_metabolite$P, method = "BH", n = n_metabolites)

# Save
res_BA_bd1_sub <- res_metabolite
```


Plot differences of bile acids by peak TMAO group

```{r}
bileacids <- c("CA", "CDCA", "DCA", "GCA", "GCDCA", "GDCA", "GLCA", "GLCAS", "GUDCA", "TCA", "TCDCA", "TDCA", "TLCA", "TMCA")
meta <- c("Row.names" ,"peakTMAO", "age", "sex", "Blood_Draw" )
combo <- c(bileacids, meta)

# Get data to graph
ggdata <- dflongPk[,colnames(dflongPk) %in% combo]

# Make blood draw numeric and showing the right relationships with regard to time
ggdata$time_numeric <- ifelse(ggdata$Blood_Draw == "BD1", 0,
                              ifelse(ggdata$Blood_Draw == "BD2", 30,
                                     ifelse(ggdata$Blood_Draw == "BD3", 180, 360)))

str(ggdata$time_numeric)

# Plot
ggplot(ggdata, aes(x = Blood_Draw, y = GCA, fill = peakTMAO)) +
  geom_boxplot()

ggplot(data = dflong, aes(x = Blood_Draw, y = GCA, fill = Blood_Draw)) +
  geom_boxplot(notch = FALSE, outlier.colour = "red") +
  xlab("Time (min)") + 
  ylab("Metabolite") +
  scale_y_continuous(trans='log2') +
  theme(text=element_text(size=13)) + 
  theme_light() +
  stat_compare_means(comparisons = list(c("0","30"), c("0", "180"), c("0","360"), c("30", "180"), c("30","360"), c("180","360")), 
  label = "p.format", method = "wilcox.test", paired=F) +
  scale_fill_brewer(palette = "RdYlBu", name = "Time (min)") #Spectral
  #ylim(0,75)

# time scale data


peak.labs <- c("Peak at 0 min", "Peak at 30 min", "Peak at 180 min", "Peak at 360 min")
names(peak.labs) <- c("peak0", "peak30", "peak180", "peak360")

# Shows individual responses
#plotData$timepoint_min <- as.numeric(levels(plotData$timepoint_min))[plotData$timepoint_min]
ggplot(ggdata, aes(x=Blood_Draw, y=GCA, group = Row.names, color = peakTMAO)) +
 # geom_rect(alpha=0.2, aes(fill = peakTMAO), xmin = -Inf,xmax = Inf,ymin = -Inf, ymax = Inf) +
  geom_line(alpha=0.4) + 
  geom_point() +
  stat_summary(aes(group = 1), geom = "point", fun = mean,
    shape = 17, size = 3, color = "blue")+
  facet_grid(peakTMAO ~ ., labeller = labeller(peakTMAO = peak.labs)) +
  xlab("Time (min)") +
  ylab(bquote("Bile Acid" ~mu~ "M")) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  #scale_color_brewer(palette = "RdYlBu") +
  scale_color_brewer(palette = "RdYlBu", name = "Peak TMAO Group", labels = c("Peak at 0 min", "Peak at 30 min", "Peak at 180 min", "Peak at 360 min")) +
  ylim(0,4)

# Show average responses by group
colnames(dflongPk2)
time_labels <- c("0 min", "30 min", "3 hr", "6 hr")

mean_data <- dplyr::group_by(dflongPk2, peakTMAO, time_numeric) %>%
             dplyr::summarise(bileacid = mean(GCA, na.rm = TRUE),
                       bileacid_SD = sd(GCA, na.rm = TRUE))
ggplot(mean_data, aes(x = time_numeric, y = bileacid, colour = peakTMAO)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(0,30,180,360),labels=time_labels) 
  #geom_errorbar(aes(x=time_numeric, ymin=bileacid-bileacid_SD, ymax=bileacid+bileacid_SD), width=0.2, alpha=0.5, size=.5)

```

Loop to see if phosphatidylcholines and other a priori metabolites change with response to the meal challenge.

```{r}
# Show average responses by group
colnames(dflongPk2)
vip <- dflongPk2[,c(91:104,216:303, 83, 339, 166, 595, 2, 1)]
n <- ncol(vip) - 2
time_labels <- c("0 min", "30 min", "3 hr", "6 hr")

# n=4 # testing

for(i in 1:n){
  metab_name <- colnames(vip)[i]
  mean_temp <- dplyr::group_by(vip, time_numeric, peakTMAO) %>%
    dplyr::summarise(metabolite = mean(!!sym(metab_name), na.rm = TRUE)) # !!sym() because dplyr is weird
              
  ggtemp <- ggplot(mean_temp, aes(x = time_numeric, y = metabolite, color = peakTMAO)) +
    geom_point() + 
    geom_line() +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    ggtitle(paste0(metab_name))
  
  ggsave(paste0("../../plots/Metab_by_Time_by_PeakTMAO/", metab_name,"_xTime_xPeakTMAO.jpg"), ggtemp,
       width = 8, 
       height = 6, 
       dpi = 300)
}
# THOSE ARE AWESOME GRAPHS

# Try adding TMAO to the plot too
vip2 <- dflongPk2[,c(49, 91:104,216:303, 83, 339, 166, 595, 2, 1)]
n <- ncol(vip2) - 2
for(i in 1:n){
  metab_name <- colnames(vip2)[i]
  
  mean_temp <- dplyr::group_by(vip2, time_numeric, peakTMAO) %>%
    dplyr::summarise(metabolite = mean(!!sym(metab_name), na.rm = TRUE)) # !!sym() because dplyr is weird
  mean_temp$metabolite_name <- metab_name
  
  mean_temp_tmao <- dplyr::group_by(vip2, time_numeric, peakTMAO) %>%
    dplyr::summarise(metabolite = mean(TMAO, na.rm = TRUE)) 
  mean_temp_tmao$metabolite_name <- "TMAO"

  # Plot  
  ggtemp_metab <- ggplot(mean_temp, aes(x = time_numeric, y = metabolite, color = peakTMAO)) +
    geom_point() + 
    geom_line() +
    theme_bw() +
    theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    scale_color_brewer(palette=c("RdYlBu"),
                    name = "Peak TMAO Group",
                    labels=c("Peak 0m", "Peak 30m", "Peak 3hr", "Peak 6hr")) +
    ylab(paste0(metab_name)) +
    xlab(NULL)
  
  ggtemp_tmao <- ggplot(mean_temp_tmao, aes(x = time_numeric, y = metabolite, color = peakTMAO)) +
    geom_point() + 
    geom_line() +
    theme_bw() +
    theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    scale_color_brewer(palette=c("RdYlBu"),
                    name = "Peak TMAO Group",
                    labels=c("Peak 0m", "Peak 30m", "Peak 3hr", "Peak 6hr")) +
    ylab("TMAO") +
    xlab(NULL)
  
  ggarrange(ggtemp_metab, ggtemp_tmao, nrow = 2, ncol=1)
  
 ggsave(paste0("../../plots/Metab_by_Time_by_PeakTMAO_facetTMAO/", metab_name,"_xTime_xPeakTMAO_facet.jpg"),
       width = 8, 
       height = 6, 
       dpi = 300)
}
# Again, awesome graphs


# Try adding SEM to the graphs
# SE
se <- function(x) sqrt(var(x)/length(x))
se(c(1,2,3,4))

vip2 <- dflongPk2[,c(49, 91:104,216:303, 83, 339, 166, 595, 2, 1)]
n <- ncol(vip2) - 2
for(i in 1:n){
  metab_name <- colnames(vip2)[i]
  
  mean_temp <- dplyr::group_by(vip2, time_numeric, peakTMAO) %>%
    dplyr::summarise(metabolite = mean(!!sym(metab_name), na.rm = TRUE),
                     metabolite_SEM = se(!!sym(metab_name))) # !!sym() because dplyr is weird
  mean_temp$metabolite_name <- metab_name
  
  mean_temp_tmao <- dplyr::group_by(vip2, time_numeric, peakTMAO) %>%
    dplyr::summarise(metabolite = mean(TMAO, na.rm = TRUE),
                     metabolite_SEM = se(TMAO)) 
  mean_temp_tmao$metabolite_name <- "TMAO"

  # Plot  
  ggtemp_metab <- ggplot(mean_temp, aes(x = time_numeric, y = metabolite, color = peakTMAO)) +
    geom_point() + 
    geom_line() +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    ylab(paste0(metab_name)) +
    xlab(NULL) +
    theme_bw() +
    theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    scale_color_brewer(palette=c("RdYlBu"),
                    name = "Peak TMAO Group",
                    labels=c("Peak 0m", "Peak 30m", "Peak 3hr", "Peak 6hr")) +
    geom_errorbar(aes(x=time_numeric, ymin=metabolite-metabolite_SEM, ymax=metabolite+metabolite_SEM), width=0.2, alpha=0.5, size=.5)
  
  ggtemp_tmao <- ggplot(mean_temp_tmao, aes(x = time_numeric, y = metabolite, color = peakTMAO)) +
    geom_point() + 
    geom_line() +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) + 
    theme_bw() +
    theme(axis.title=element_text(size=14),
        axis.text=element_text(size=12),
        legend.title=element_text(size=14),
        legend.text=element_text(size=12)) +
    scale_x_continuous(breaks = c(0,30,180,360), labels = time_labels) +
    scale_color_brewer(palette=c("RdYlBu"),
                    name = "Peak TMAO Group",
                    labels=c("Peak 0m", "Peak 30m", "Peak 3hr", "Peak 6hr")) +
    ylab("TMAO") +
    xlab(NULL) +
    geom_errorbar(aes(x=time_numeric, ymin=metabolite-metabolite_SEM, ymax=metabolite+metabolite_SEM), width=0.2, alpha=0.5, size=.5)
  
  ggarrange(ggtemp_metab, ggtemp_tmao, nrow = 2, ncol=1)
  
 ggsave(paste0("../../plots/Metab_by_Time_by_PeakTMAO_facetTMAO_sem/", metab_name,"_xTime_xPeakTMAO_facet.jpg"),
       width = 9, 
       height = 6, 
       dpi = 300)
}

```


